---
title: "BiocPropagate"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Propagation Checks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The purpose of this package is to consolidate propagation decisions given
the package build results in R Universe and Bioconductor propagation criteria.

## Checking propagation

`check_propagation(args)` where `args` is a named list passing

| Element | Required | Description |
|---|---|---|
| `package` | x | Package name |
| `universe` | x | r-universe universe name |
| `jobs` | x | A data.frame of r-universe job data |
| `source_path` | x | Path to the extracted source package |
| `criteria` | | A criteria list, defaults to `default_criteria()` |

The `universe` determines the Bioconductor branch internally. Bioconductor's
published package data is always fetched live.

## Design

The package has a default set of propagation criteria but can allow additional
criteria to be specified or excluded. There are two types of criteria operating
at different levels. Unlike platform criteria, such as `platform_status` that
independently evaluate the `R CMD check` result of the package on a platform,
gate criteria, such as `valid_version` or `no_remotes`, check for package-level
problems that render the propagation as `FALSE` for all platforms.

```
GATES (once, package-level)
  vignettes, version, no_large_files, no_remotes,
  no_secrets, no_merge_conflicts
        |
   any fail? -> every row's propagate = FALSE
        |
   all pass
        v
PLATFORM CRITERIA (each _jobs row, independently)
  status, unsupported, platform_version
        |
   all pass -> that row's propagate = TRUE
```

The result of `check_propagation` is the `_jobs` data.frame with an extra
`propagate` column that is `TRUE` or `FALSE`.

### Details

- Unsupported platforms in `Config/Bioconductor/UnsupportedPlatforms` evaluate
to `FALSE`
- In R Universe, vignettes are checked in the `source` `_job`
- The package version for a platform is evaluated by the `platform_version`
criteria to prevent a version from being propagated again
- OS-r_version-arch combinations not valid for a given Bioconductor branch
evaluate to `NA`

#### Default critera

```{r default_critera}
library(BiocPropagate)

criteria <- default_criteria()
names(criteria$gates)
names(criteria$platform)
```

#### Additional critera

Additional criteria may be added. Gates may also be removed.

```{r bioc-checks, eval = FALSE}
criteria <- default_criteria()

# Add
criteria <- register_criterion(
    criteria, "always_pass",
    function(pkg_data, branch, bioc_pkg_data, source_path)
        list(pass = TRUE, message = NA_character_),
        type = "gates"
    )
names(criteria$gates)

# Remove
criteria <- unregister_gates(criteria, "always_pass")
names(criteria$gates)
```

Packages granted exemptions from gate-level criteria can also be defined
in an exemptions.txt DCF file in the package manifest with the following format:

```
# manifest/exemptions.txt
Package: package1
Exemptions: no_large_files, no_remotes

Package: package2
Exemptions: no_remotes
```

Note: A deploy key should be provided to the organization building the universe
if the manifest is private.

## Example

We'll do an example with live `_jobs` data from r-universe's public JSON API. 
`DESCRIPTION` is reconstructed from the same payload and the result will be
based on the default criteria.

```{r example}
library(BiocPropagate)

pkg <- jsonlite::fromJSON("https://bioc-release.r-universe.dev/BiocCheck/json")

dir <- tempfile()
dir.create(dir)
writeLines(
    c(paste("Package:", pkg[["Package"]]), paste("Version:", pkg[["Version"]])),
    file.path(dir, "DESCRIPTION")
)

args <- list(
    package = pkg[["Package"]],
    universe = "bioc-release",
    jobs = pkg[["_jobs"]],
    source_path = dir
)

propagation <- check_propagation(args)
print(propagation)
```
